---
title: "Big Pick Energy Baseball"
author: "Patrick Milum"
date: "2023-05-10"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

# Not meant to knit! Don't do it. It will take you forever. Just an organized way to keep things updated.

### [**Daily Runs**]{.underline}

#### Load packages, get colors/logos

For adding colors/logos to a plot: `left_join` colors2 data frame with
whatever data you want to visualize. See final step here.

```{r include = FALSE}
library(tidyverse)
library(ggimage)
library(teamcolors)
library(ggpubr)
library(baseballr)
library(ggrepel)
library(lubridate)
library(ggplot2)
library(rvest)
library(skimr)
library(writexl)

colors <- teamcolors %>% filter(league == "mlb") %>%
  mutate(name = case_when(
    name == "Cleveland Indians" ~ "Cleveland Guardians", 
    TRUE ~ name)) %>%
  mutate(mascot = case_when(
    mascot == "Indians" ~ "Guardians", 
    TRUE ~ mascot)) %>%
  mutate(sportslogos_name = case_when(
    sportslogos_name == 'Cleveland Indians' ~ 'Cleveland Guardians', 
    TRUE ~ sportslogos_name))

colors2 <- colors %>%
  mutate(team.abbr = case_when(mascot == 'Diamondbacks' ~ 'ARI', 
                               mascot == 'Braves' ~ 'ATL', 
                               mascot == 'Orioles' ~ 'BAL', 
                               mascot == 'Red Sox' ~ 'BOS', 
                               mascot == 'Cubs' ~ 'CHC', 
                               mascot ==  'White Sox' ~ 'CHW', 
                               mascot ==  'Reds' ~ 'CIN', 
                               mascot == 'Indians' ~ 'CLE', 
                               mascot == 'Rockies' ~ 'COL', 
                               mascot == 'Tigers' ~ 'DET', 
                               mascot == 'Astros' ~ 'HOU', 
                               mascot == 'Royals' ~ 'KCR', 
                               mascot == 'Angels' ~ 'LAA', 
                               mascot == 'Dodgers' ~ 'LAD', 
                               mascot == 'Marlins' ~ 'MIA', 
                               mascot == 'Brewers' ~ 'MIL', 
                               mascot == 'Twins' ~ 'MIN', 
                               mascot == 'Mets' ~ 'NYM', 
                               mascot == 'Yankees' ~ 'NYY', 
                               mascot == 'Athletics' ~ 'OAK', 
                               mascot == 'Phillies' ~ 'PHI', 
                               mascot == 'Pirates' ~ 'PIT', 
                               mascot == 'Padres' ~ 'SDP', 
                               mascot == 'Giants' ~ 'SFG', 
                               mascot == 'Mariners' ~ 'SEA', 
                               mascot == 'Cardinals' ~ 'STL', 
                               mascot == 'Rays' ~ 'TBR', 
                               mascot == 'Rangers' ~ 'TEX', 
                               mascot == 'Blue Jays' ~ 'TOR', 
                               mascot == 'Nationals' ~ 'WSN', 
                               mascot == 'Guardians' ~ 'CLE'))

team.batting.plot <- left_join(team.batting, colors2, 
                               join_by(Team == team.abbr))
team.pitching.plot <- left_join(team.pitching, colors2, 
                                join_by(Team == team.abbr))
```

#### Scrape pitch by pitch data from Statcast

This should be done in (usually) 5-6 day pulls. Once the scrape has
reached its limit, you'll need to end the date and create a new call.
You do NOT want to call the previous scrapes once you have them in your
Global Environment. They take a while.

```{r}
#scrape1.pbp <- statcast_search("2023-03-30", "2023-04-05", player_type = "pitcher") 
#scrape2.pbp <- statcast_search("2023-04-06", "2023-04-10", player_type = "pitcher")
#scrape3.pbp <- statcast_search("2023-04-11", "2023-04-17", player_type = "pitcher") 
#scrape4.pbp <- statcast_search("2023-04-18", "2023-04-23", player_type = "pitcher")
#scrape5.pbp <- statcast_search("2023-04-24", "2023-04-30", player_type = "pitcher")
#scrape6.pbp <- statcast_search("2023-05-01", "2023-05-06", player_type = "pitcher")
#scrape7.pbp <- statcast_search("2023-05-07", "2023-05-13", player_type = "pitcher")
scrape8.pbp <- statcast_search("2023-05-14", today(), player_type = "pitcher")


```

From here you'll need to bind the scrapes together into one data frame.
Don't forget to add the new scrapes here! You'll also create a data
frame from the Chadwick Database which has the names and player IDs for
all MLB players.

```{r include = FALSE}
season <- rbind(scrape1.pbp, scrape2.pbp, scrape3.pbp, scrape4.pbp, scrape5.pbp, scrape6.pbp, 
                scrape7.pbp, scrape8.pbp) 
#chad <- chadwick_player_lu()%>%
#  select(key_mlbam, name_last, name_first) %>%
 # filter(!is.na(key_mlbam)) 
```

Finally, you can left_join your 'season' data frame with 'chad' and
sort, select, arrange, and mutate your variables! This is the cool part.

```{r}
season.pbp <- left_join(season, chad, join_by("batter" == "key_mlbam")) %>%
  rename(., "batter_id" = "batter") %>%
  rename(., "pitcher_id" = "pitcher") %>%
  rename(., "pitcher" = "player_name") %>%
  unite(., batter, c(name_last, name_first)) %>%
  select(game_date, pitcher, p_throws, batter, stand, pitch_type, pitch_name, pitch_number, balls, strikes, outs_when_up, at_bat_number, 
         release_speed, release_spin_rate, release_extension, release_pos_x, release_pos_z, description, events, des, home_team, away_team, inning, 
         inning_topbot, launch_speed, launch_angle, estimated_ba_using_speedangle, home_score, away_score, post_home_score, post_away_score, 
         bat_score, fld_score, post_bat_score, post_fld_score, game_pk, at_bat_number, pitcher_id, batter_id, everything()) %>%
  arrange(desc(game_date), game_pk) %>%
  mutate(batter = str_replace_all(batter, "_", ", ")) %>%
  mutate(pitcher.home = ifelse(inning_topbot == "Top", "Home", "Away")) %>%
  mutate(pitcher.team = ifelse(pitcher.home == "Home", paste(home_team), paste(away_team))) %>%
  mutate(batter.home = ifelse(inning_topbot == "Bot", "Home", "Away")) %>%
  mutate(batter.team = ifelse(batter.home == "Home", paste(home_team), paste(away_team))) %>%
  mutate(batter.team = case_when(batter.team == 'AZ' ~ 'ARI', 
                          batter.team == 'KC' ~ 'KCR', 
                          batter.team == 'SF' ~ 'SFG', 
                          batter.team == 'WSH' ~ 'WSN', 
                          batter.team == 'SD' ~ 'SDP', 
                          batter.team == 'TB' ~ 'TBR', 
                          batter.team == 'CWS' ~ 'CHW', 
                          TRUE ~ batter.team)) %>%
  mutate(pitcher.team = case_when(pitcher.team == 'AZ' ~ 'ARI', 
                          pitcher.team == 'KC' ~ 'KCR', 
                          pitcher.team == 'SF' ~ 'SFG', 
                          pitcher.team == 'WSH' ~ 'WSN', 
                          pitcher.team == 'SD' ~ 'SDP', 
                          pitcher.team == 'TB' ~ 'TBR', 
                          pitcher.team == 'CWS' ~ 'CHW', 
                          TRUE ~ pitcher.team)) %>%
  mutate(home_team = case_when(home_team == 'AZ' ~ 'ARI', 
                          home_team == 'KC' ~ 'KCR', 
                          home_team == 'SF' ~ 'SFG', 
                          home_team == 'WSH' ~ 'WSN', 
                          home_team == 'SD' ~ 'SDP', 
                          home_team == 'TB' ~ 'TBR', 
                          home_team == 'CWS' ~ 'CHW', 
                          TRUE ~ home_team)) %>%
  mutate(away_team = case_when(away_team == 'AZ' ~ 'ARI', 
                          away_team == 'KC' ~ 'KCR', 
                          away_team == 'SF' ~ 'SFG', 
                          away_team == 'WSH' ~ 'WSN', 
                          away_team == 'SD' ~ 'SDP', 
                          away_team == 'TB' ~ 'TBR', 
                          away_team == 'CWS' ~ 'CHW', 
                          TRUE ~ away_team)) 
```

#### Individual player stats (season long)

This comes from Fangraphs, which can be different from Statcast
sometimes (just a head's up). Pitching has 303 variables, qual = IP.
Batting has 289 variables, qual = AB.

```{r}
pitchers <- fg_pitch_leaders(2023, 2023, qual = .1) %>%
  mutate(Pitches.per.PA = round((Pitches/TBF), 2)) %>%
  select(playerid,Name, Team, G, GS, W, L, ER, ERA, IP, Relief_IP, LOB_pct, SV, BS, TBF, H, R, RS, RS_9, HR, BB, SO, Balls, Strikes, Pitches, 
         Pitches.per.PA, Zone_pct, BB_pct, BB_9, K_pct, K_9, H_9, HR_9, AVG, WHIP, BABIP, Pull_pct, Cent_pct, Oppo_pct,
               GB_pct, FB_pct, LD_pct, WPA, RE24, FBall_pct, FBv, 
               SL_pct, SLv, CT_pct, CTv, CB_pct, CBv, CH_pct, CHv, Swing_pct, `O-Swing_pct`, `Z-Swing_pct`, 
               Contact_pct, `O-Contact_pct`, `Z-Contact_pct`, `F-Strike_pct`, SwStr_pct, Hard_pct, Pace, everything())

batters <- fg_batter_leaders(2023, 2023, qual = 1) %>%
  mutate(Pitches.per.PA = round((Pitches/PA), 2), 
         R.per.PA = round((R/PA), 2))%>%
  select(playerid, Name, Team, G, AB, PA, H, `1B`, `2B`, `3B`, HR, R, R.per.PA, Pitches, 
         Balls, Strikes, Pitches.per.PA, RBI, BB, SO, GDP, SB, CS, AVG, OBP, SLG, OPS, 
          BABIP, BB_pct, K_pct, GB_pct, FB_pct, wOBA, wRC, WAR, WPA, RE24, FBall_pct, FBv, 
          SL_pct, SLv, CT_pct, CTv, CB_pct, CBv, CH_pct, CHv, Swing_pct, `O-Swing_pct`, `Z-Swing_pct`, 
          Contact_pct, `O-Contact_pct`, `Z-Contact_pct`, `F-Strike_pct`, SwStr_pct, Hard_pct, Pace, everything())

```

#### Team stats (season long)

This also comes from Fangraphs. After you mutate and select some key
variables, you calculate the mean for every stat so you can have a basis
for comparison. You'll then bind the mean row with the data frame.

```{r include = FALSE}
team.batting <- fg_team_batter(2023, 2023) %>%
  mutate(Pitches.per.AB = round((Pitches/AB), 3), 
         R.per.AB = round((R/AB), 2), 
         R.per.PA = round((R/PA), 2)) %>%
  select(Team, AB, PA, H, R, R.per.AB, R.per.PA, HR, RBI, BB, SO, GDP, SB, CS, AVG, OBP, SLG, OPS, 
         BABIP, Pitches, Pitches.per.AB, BB_pct, K_pct, GB_pct, FB_pct, wOBA, wRC, WAR, WPA, RE24, FBall_pct, FBv, 
         SL_pct, SLv, CT_pct, CTv, CB_pct, CBv, CH_pct, CHv, Swing_pct, `O-Swing_pct`, `Z-Swing_pct`, 
         Contact_pct, `O-Contact_pct`, `Z-Contact_pct`, `F-Strike_pct`, SwStr_pct, Hard_pct, Pace, everything())
bat.mean <- summarize_all(team.batting, mean, 2)
team.batting <- rbind(team.batting, bat.mean)
team.batting[31, 1] <- "Mean"


team.pitching <- fg_team_pitcher(2023, 2023) %>%
  mutate(pitches.per.PA = round(Pitches/TBF, 2), 
         G = W + L, 
         RA.per.G = round((R/G), 2), 
         RD = RS - R, 
         RD.per.G = round(RD/G, 2)) %>%
  select(Team, G, W, L, ERA, IP, Relief_IP, SV, BS, TBF, H, R, RS, RS_9, RA.per.G, HR, BB, SO, Balls, Strikes, Pitches, pitches.per.PA, BB_pct, BB_9, K_pct, K_9, H_9, HR_9, AVG, WHIP, BABIP, 
         GB_pct, FB_pct, WPA, RE24, FBall_pct, FBv, 
         SL_pct, SLv, CT_pct, CTv, CB_pct, CBv, CH_pct, CHv, Swing_pct, `O-Swing_pct`, `Z-Swing_pct`, 
         Contact_pct, `O-Contact_pct`, `Z-Contact_pct`, `F-Strike_pct`, SwStr_pct, Hard_pct, Pace, everything())
pitch.mean <- summarize_all(team.pitching, mean, 2)
team.pitching <- rbind(team.pitching, pitch.mean)
team.pitching[31,1] <- "Mean"
```

------------------------------------------------------------------------

### [**Pitching Finds**]{.underline}

Here we are going to dive into the data that we've been collecting for
the 2023 season. Make sure to identify whether each data frame, metric,
new variable, etc. are measuring pitching performance or batting
performance. This should be growing all the time (ideally). Make sure to
provide context and detail for each code. What does it measure, what
needs to be considered, are there outliers, etc.? Let's go!

#### Total Pitches by Type (arsenal)

There is probably a much easier way to organize this (in fewer lines of
code), but this tells the number of pitches each pitcher has thrown over
the course of a season. Not much to infer from this, just interesting to
see the different pitch arsenals. Minimum of 20 total pitches thrown
(this will be raised as the year progresses).

```{r include = FALSE}
arsenal <- season.pbp %>%
  group_by(pitcher, pitcher.team) %>%
  summarize(total.pitches = n(), 
            `4-Seam Fastball` = sum(pitch_type == 'FF'), 
            ff.pct = round(`4-Seam Fastball`/total.pitches, 2), 
            `Curveball` = sum(pitch_type == 'CU'), 
            cu.pct = round(Curveball/total.pitches, 2), 
            `Changeup` = sum(pitch_type == 'CH'), 
            ch.pct = round(Changeup/total.pitches, 2), 
            `Slider` = sum(pitch_type == 'SL'), 
            sl.pct = round(Slider/total.pitches, 2), 
            `Cutter` = sum(pitch_type == 'FC'), 
            fc.pct = round(Cutter/total.pitches, 2), 
            `Sinker` = sum(pitch_type == 'SI'), 
            si.pct = round(Sinker/total.pitches, 2), 
            `Split-Finger` = sum(pitch_type == 'FS'), 
            fs.pct = round(`Split-Finger`/total.pitches, 2), 
            `Knuckle-Curve` = sum(pitch_type == 'KC'), 
            kc.pct = round(`Knuckle-Curve`/total.pitches, 2), 
            `Sweeper` = sum(pitch_type == 'ST'), 
            st.pct = round(Sweeper/total.pitches, 2), 
            `Slurve` = sum(pitch_type == 'SV'), 
            sv.pct = round(Slurve/total.pitches, 2), 
            `Screwball` = sum(pitch_type == 'SC'), 
            sc.pct = round(Screwball/total.pitches, 2), 
            `Forkball` = sum(pitch_type == 'FO'), 
            fo.pct = round(Forkball/total.pitches, 2), 
            `Other` = sum(pitch_type == 'FA'), 
            other.pct = round(Other/total.pitches, 2)) %>%
  filter(total.pitches >= 20) %>%
  arrange(desc(total.pitches))
```

#### Details of Pitch Types (pitch.details)

This call is going to give you the average velocity, spin, extension of
each pitch type. It will also tell you how many were thrown for balls,
strikes, ball-fours, strike-threes, and homers. One way to improve on
this would be to add movement profiles for each pitch. This would help
in the future with visualizing the data. Note: the strike.pct column
tells if the pitch was a strike OR made contact with (including put in
play). Minimum of 20 tosses of that particular pitch. Pairs well with
batter.hits.by.type.

```{r include = FALSE}
pitch.details <- season.pbp %>%
  group_by(pitcher, pitch_name, pitcher.team) %>%
  summarize(count = n(),
            avg.speed = round(mean(release_speed, na.rm = T), 2), 
            avg.spin = round(mean(release_spin_rate, na.rm = T), 2), 
            avg.ext = round(mean(release_extension, na.rm = T), 2), 
            hits = sum(events == 'single' | events == 'double' | events == 'triple' | events == 'home_run'), 
            balls = sum(type == "B"), 
            strikes = sum(type == "S" | type == "X"), 
            ball.fours = sum(events == 'walk'), 
            strike.threes = sum(events == 'strikeout'), 
            homeruns = sum(events == 'home_run'), 
            hr.freq = round(homeruns/count, 2),
            hit.pct = round(hits/count, 2)) %>%
  mutate(strike.pct = round(strikes/count, 2)) %>%
  filter(count >= 15)
```

#### First Five Innings, Pitchers (f5.pitcher.logs, f5.pitcher.season) \*work on visualization

Here we have some information about how each starting pitcher does in
their first five innings of work. This is mainly used for F5 betting
purposes, although it can help understand who is struggling early and
who starts well. The first call (`first.five.pitchers`) gives game logs.
This can help look at outliers and see how a pitcher matched up against
a particular team. From there, we create the `first.five` data frame,
which shows us a pitcher's season totals. #ggplot viz needs work#

```{r }
f5.pitcher.logs <- season.pbp %>%
  filter(inning <= 5) %>%
  group_by(pitcher, game_pk, pitcher.team, batter.team) %>%
  summarize(PA = (max(at_bat_number) - min(at_bat_number) + 1),
            hits = sum(events %in% c('single', 'double', 'triple', 'home_run')), 
            walks.hbp.error = sum(events %in% c('walk', 'hit_by_pitch', 'catcher_interf', 'field_error')), 
            total.batters.reached = hits + walks.hbp.error, 
            pct.batters.reached = round(total.batters.reached/PA, 2), 
            runs.allowed = sum(post_bat_score - bat_score), 
            pct.runners.scored = round(runs.allowed/total.batters.reached, 2))

f5.pitcher.season <- f5.pitcher.logs %>%
  group_by(pitcher, pitcher.team) %>%
  summarize(PA = sum(PA), 
            hits = sum(hits), 
            walks.hbp.error = sum(walks.hbp.error), 
            total.batters.reached = sum(total.batters.reached), 
            pct.batters.reached = round(mean(pct.batters.reached), 2), 
            runs.allowed = sum(runs.allowed), 
            pct.runners.scored = round(mean(pct.runners.scored), 2), 
            runs.per.PA= round(mean(runs.allowed/PA), 2)) %>%
  arrange(desc(runs.per.PA))
f5.pitcher.season <- f5.pitcher.season %>%
  filter(PA >= 30)

#ggplot(f5.season, aes(pct.batters.reached, pct.runners.scored)) +
  #geom_point(cex = f5.season$runs.per.PA*20) + 
  #geom_text_repel(label = paste(f5.season$pitcher, f5.season$PA)) +
  #geom_smooth(method = 'lm')
```

#### First Five Innings, Pitcher v. Batter Matchups (f5.AB)

This is every AB that has occurred during the first five innings of
games. This will tell you what the score was (for both teams and
combined), the outcome of the AB, and whether there was a CS/GIDP in the
AB. There seems to be something missing here, but I can't quite figure
out how to improve it (or make it useful).

```{r include = FALSE}
f5.AB <- season.pbp %>%
  filter(inning <= 5) %>%
  group_by(home_team, away_team, batter, game_pk, game_date, at_bat_number, pitcher) %>%
  summarize(home.score = max(home_score), 
            away.score = max(away_score), 
            total.runs = home.score + away.score, 
            pitches = max(pitch_number), 
            hits = sum(events %in% c('single', 'double', 'triple', 'home_run')), 
            walks.hbp.error = sum(events %in% c('walk', 'hit_by_pitch', 'catcher_interf', 'field_error')), 
            CS.pickoff = sum(events %in% c('caught_stealing_home', 'pickoff_1b', 'caught_stealing_2b', 'caught_stealing_3b')),
            double.play = sum(events %in% c('double_play', 'strikeout_double_play', 'sac_fly_double_play', 'grounded_into_double_play'))) %>%
  select(game_date, game_pk, pitcher, everything()) %>%
  arrange(game_pk, at_bat_number)
```

#### Pitcher v. Batter Matchups: Outcomes (matchup.outcomes)

How many times did a pitcher and batter match up and what happened each
time? Gotcha. The key here is grouping twice, excluding the team names
in the second call, and left joining the two data frames.

```{r include = FALSE}
outcomes <- season.pbp %>%
  filter(events != '') %>%
  group_by(batter, batter.team, pitcher, pitcher.team, description, events) %>%
  summarize(rate = n()) %>%
  ungroup()
new.outcome <- outcomes %>%
  group_by(batter, pitcher) %>%
  summarize(PAs = sum(rate))
matchup.outcomes <- outcomes %>%
  left_join(new.outcome, join_by('batter' == 'batter', 'pitcher' == 'pitcher')) %>%
  mutate(outcome.pct = round(rate/PAs, 2))

```

#### Pitcher Game Logs using built function

This will pull a pitcher's game logs and bind them with the game logs of
the opposing team's starting pitcher. I've chosen some key stats but
have left out the pitch usage. This could be important when diagnosing
matchups. (I suppose a beneficial adjustment could be to bind a
pitcher's stats with the opposing team batting stats.) Note: must call
function first (found in Functions section below).

```{r include = FALSE}
Corbin <- pitcher.game.logs("Corbin", "Patrick", 2023)
Taillon <- pitcher.game.logs("Taillon", "Jameson", 2023)
game1 <- rbind(Corbin, Taillon, fill = T) %>%
  select(Team, PlayerName, Date, Opp, W, L, IP, TBF, Balls, Strikes, Pitches, H, R, ER, ERA, 
         HR, BB, HBP, SO, `K%`, `K/9`, `BB/9`, `BB%`, `H/9`, `HR/9`, GB, FB, LD, AVG, WHIP, `LOB%`, 
         RS, `RS/9`, EV, `HardHit%`, Barrels, `Barrel%`)
#write_xlsx(game1, "WshChcThu.xlsx")
```

------------------------------------------------------------------------

### [Batting Finds]{.underline}

#### Batter Hits by Type (batter.hits.by.type)

This tells us how each batter does against each pitch type. Strikes are
true strikes, not balls put in play (in play is its own variable here).
This is good to see how a batter does against a pitch that their
opponent throws well and often. Pairs well with pitch.details.

```{r include = FALSE}
batter.hits.by.type <- season.pbp %>%
  group_by(batter, pitch_name) %>%
  summarize(singles = sum(events == 'single'), 
            doubles = sum(events == 'double'), 
            triples = sum(events == 'triple'), 
            home_runs = sum(events == 'home_run'), 
            in.play = sum(type == 'X'), 
            hits = sum(events %in% c('single', 'double', 'triple', 'home_run')),  
            strikes = sum(type == "S"), 
            called.strike = sum(description == 'called_strike'), 
            balls = sum(type == "B"), 
            pitches = n(), 
            b.pct = round((balls/pitches), 3),
            strike.pct = round((strikes/pitches), 3),
            in.play.pct = round(in.play/pitches, 3),  
            avg.pitch.speed = round(mean(release_speed), 2), 
            hit.pct = round(hits/pitches, 2)) %>%
  mutate(total.pitches = sum(pitches), 
         pct = round(pitches/total.pitches, 3)) %>%
  select(batter, pitch_name, pitches, in.play, avg.pitch.speed, hit.pct, hits, singles, doubles, triples, home_runs, strikes, called.strike, balls, 
         strike.pct, b.pct, in.play.pct, total.pitches, pct) %>%
  arrange(desc(in.play))
```

#### Go Ahead RBIs (go.ahead.rbi)

This measures each time that a player got a hit, sacrifice, or fielder's
choice which lead to a go ahead run while their team was either down or
tied. This can be filtered down to look at certain innings (maybe 7th or
later). This will avoid the 1st inning 0-0 'go ahead RBI.'

```{r}
go.ahead.rbi <- season.pbp %>%
  filter(post_bat_score > bat_score & post_bat_score > post_fld_score & fld_score >= bat_score) %>%
  mutate(go.ahead.rbi = post_bat_score - bat_score) %>%
  group_by(batter) %>%
  summarize(hits = sum(events %in% c('single', 'double', 'triple', 'home_run')),
            sac = sum(events %in% c('sac_bunt', 'sac_fly')), 
            fc = sum(events %in% c('fielders_choice', 'fielders_choice_out')), 
            rbi = sum(go.ahead.rbi))
```

------------------------------------------------------------------------

### [**Betting Lines**]{.underline}

#### Over/Under Records by Team (over.under)

This scrapes from the TeamRankings website for betting. Something that I
need to figure out is how to automate this into a spreadsheet that will
update daily, instead of saving as .xlsx file and importing it into
Excel. Be sure that your working directory is set before writing the
file.

```{r}
over.under <- read_html("https://www.teamrankings.com/mlb/trends/ou_trends/") %>%
  html_table(fill = T) %>%
  as.data.frame() %>%
  rename(., "Total" = "Total....") %>%
  rename(., "Under" = "Under..") %>%
  rename(., "Over" = "Over..") 
write_xlsx(over.under, "OverUnderTotals.xlsx")
```

#### Team Game Logs and F5 (team.game.logs & f5.game.logs)

I've gathered game by game scores and (in a separate data frame) the
same info for the first 5 innings. You need to `left_join` these
together. In all, it is two scrapes and a join, then aggregate more for
deeper analysis. Team Game Logs would be so good with an O/U line for
each game. How to bring that in...

```{r include = FALSE}
team.game.logs <- season.pbp %>%
  group_by(game_pk, game_date, home_team, away_team) %>%
  summarize(runs.h = max(post_home_score), 
            runs.a = max(away_score), 
            total.runs = runs.h + runs.a, 
            score = paste(runs.h, "-", runs.a), 
            innings = max(inning), 
            hits = sum(events %in% c('single', 'double', 'triple', 'home_run'))) %>%
  mutate(win= if_else(runs.h > runs.a, 1, -1),
    winner = if_else(runs.h > runs.a,  "Home", "Away")) %>%
  arrange(desc(game_date)) 

f5 <- season.pbp %>%
  filter(inning <= 5) %>%
  group_by(game_pk, game_date, home_team, away_team) %>%
  summarize(`f5.runs.h` = max(post_home_score), 
            `f5.runs.a` = max(post_away_score), 
            `f5.total.runs` = `f5.runs.h` + `f5.runs.a`)

f5.game.logs <- team.game.logs %>%
  left_join(f5, by = c('game_pk', 'game_date', 'home_team', 'away_team')) %>%
  mutate(`%runs.h.f5` = round(`f5.runs.h`/runs.h, 2), 
            `%runs.a.f5` = round(`f5.runs.a`/runs.a, 2),
            `%total.runs.f5` = round(`f5.total.runs`/total.runs, 2)) %>%
  select(game_date, game_pk, innings, home_team, away_team, runs.h, runs.a, total.runs, 
         `f5.runs.h`, `f5.runs.a`, `f5.total.runs`, `%runs.h.f5`, `%runs.a.f5`, 
         `%total.runs.f5`, hits, score)
```

#### Home/Away Splits Scoring Over Season (home.five, away.five)

What does each team score (and give up) as home team AND away team, both
in total and in the first five? The previoius build (team.game.logs)
show info for each game, this is overall for the season. Should find a
way to merge these but I kind of like them separate. May be able to call
both of these in a matchup function...good idea. This is a subset of
team.game.logs

```{r include = FALSE}
home.five <- f5.game.logs %>%
       group_by(home_team) %>%
       summarize(`%runs.h.f5` = round(mean(`%runs.h.f5`, na.rm = TRUE), 2), 
                 total.runs.home = sum(runs.h), 
                 f5.runs.h = sum(`f5.runs.h`), 
                 f5.runs.a = sum(`f5.runs.a`), 
                 `%runs.a.f5` = round(mean(`%runs.a.f5`, na.rm = TRUE), 2), 
                 games = n(), 
                 f5.run.diff = f5.runs.h - f5.runs.a, 
                 RS.per.5 = round(f5.runs.h/games, 2), 
                 RA.per.5 = round(f5.runs.a/games, 2), 
                 runs.per = RS.per.5 + RA.per.5)

away.five <- f5.game.logs %>%
  group_by(away_team) %>%
  summarize(`%runs.a.f5` = round(mean(`%runs.a.f5`, na.rm = TRUE), 2), 
            total.runs.away = sum(runs.a), 
            f5.runs.a = sum(`f5.runs.a`), 
            f5.runs.h = sum(`f5.runs.h`), 
            `%runs.h.f5` = round(mean(`%runs.h.f5`, na.rm = TRUE), 2), 
            games = n(), 
            f5.run.diff = f5.runs.a - f5.runs.h, 
            RS.per.5 = round(f5.runs.a/games, 2), 
            RA.per.5 = round(f5.runs.h/games, 2), 
            runs.per = RS.per.5 + RA.per.5)

```

#### No Runs First Inning (NRFI)

This took some work, conceptually. Aggregating was simple but it was a
challenge to get the total runs and not runs for every game. Turns out
you have to subtract the score from the post_score. Still confused about
how it works, but it works!

```{r include = FALSE}
home.NRFI <- season.pbp %>%
  filter(inning == 1) %>%
  group_by(home_team) %>%
  summarize(RS.home = sum(post_home_score - home_score),
            RA.home = sum(post_away_score - away_score), 
            home.games = length(unique(game_pk)), 
            RS.per.home = round(RS.home/home.games, 2), 
            RA.per.home = round(RA.home/home.games, 2))


away.NFRI <- season.pbp %>%
  filter(inning == 1) %>%
  group_by(away_team) %>%
  summarize(RS.away = sum(post_away_score - away_score),
            RA.away = sum(post_home_score - home_score), 
            away.games = length(unique(game_pk)), 
            RS.per.away = round(RS.away/away.games, 2), 
            RA.per.away = round(RA.away/away.games, 2))

NRFI <- home.NRFI %>%
  left_join(away.NFRI, join_by(home_team == away_team)) %>%
  mutate(home.game.runs = RS.home + RA.home, 
         away.game.runs = RS.away + RA.away) %>%
  rename("team" = "home_team")
```

#### Combined daily matchups, strength rating, and predicted runs (today.games)

This code is UGLY. There is definitely a better way to
consolidate/streamline this code but for now...

```{r}
#get game pks, join with colors2 for abbreviations, rename appropriate teams and select key variables
today.games <- get_game_pks_mlb(today(), level_ids = c(1)) %>%
  select(game_pk, officialDate, dayNight, seriesGameNumber, teams.away.team.name, teams.home.team.name)
today.games <- today.games %>%
  left_join(colors2, join_by('teams.away.team.name' == 'name')) %>%
  left_join(colors2, join_by('teams.home.team.name' == 'name'))
today.games <- today.games %>%
  rename('away_team' = 'team.abbr.x') %>%
  rename('home_team' = 'team.abbr.y') %>%
  select(game_pk, officialDate, away_team, home_team, officialDate) 
#add strengths for away team and home team, aggregate to find prediction of game
today.games <- today.games %>%
  left_join(strength.away, by = c('away_team' = 'team'))
today.games <- today.games %>%
  left_join(strength.home, by = c('home_team' = 'team')) 
today.games <- today.games %>%
  mutate(prediction = away.strength + home.strength, 
         absolute = abs(prediction), 
         conf = rank(absolute))
#use team.game.logs to aggregate RS and RA per game for both home and away team in their own dataframes
away.stats <- team.game.logs %>%
  select(-home_team) %>%
  group_by(away_team) %>%
  summarize(away.R.per.game = sum(runs.a)/n(), 
            away.RA.per.game = sum(runs.h)/n()) 
home.stats <- team.game.logs %>%
  select(-away_team) %>%
  group_by(home_team) %>%
  summarize(home.R.per.game = sum(runs.h)/n(), 
            home.RA.per.game = sum(runs.a)/n()) %>%
  rename('team' = 'home_team')
#join the RA and RS dataframse to the day's game slate, aggregate a final score prediction
today.games <- today.games %>%
  left_join(away.stats, by = 'away_team') %>%
  left_join(home.stats, by = c('home_team' = 'team')) %>%
  mutate(pred.total = (away.R.per.game + away.RA.per.game + home.R.per.game + home.RA.per.game)/2)
today.games <- today.games %>%
  select(away_team, home_team, prediction, pred.total, conf, everything())


```

#### First attempt at binding daily predictions with outcomes (dummy, results, year.pred.results)

I believe that I have to initially add the game slate and predictions to
`year.pred.results` and from now on, bind it as follows. To reset and
create new mutate variables, year.pred.results \<- results.

This is fucked. UNLESS you remove duplicated rows, which is a good way
to end this call.

```{r}
dummy <- today.games
results <- merge(dummy, team.game.logs, by = 'game_pk')
year.pred.results <- 
  rbind(year.pred.results, results, fill = TRUE) %>%
  mutate(over.under = if_else(pred.total > total.runs, "Under", "Over"), 
         unders = if_else(pred.total > total.runs, 1, 0), 
         overs = if_else(pred.total < total.runs, 1, 0), 
         winner.check = case_when(winner == 'Home' & prediction >= 0 ~ 1, 
                                    winner == 'Away' & prediction < 0 ~ 1, 
                                    winner == 'Home' & prediction < 0 ~ 0, 
                                    winner == 'Away' & prediction >= 0 ~ 0), 
         pred.diff = pred.total - total.runs) 
year.pred.results <- year.pred.results[!duplicated(year.pred.results)]
mean(year.pred.results$winner.check)
```

------------------------------------------------------------------------

#### Average margin of win/loss for home/away splits

It is exactly what it sounds like it is.

```{r}
margins.home.wins <- team.game.logs %>%
  filter(winner == 'Home') %>%
  group_by(home_team) %>%
  summarize(home.w.marg = mean(runs.h - runs.a)) %>%
  rename('team' = 'home_team')
margins.home.loss <- team.game.logs %>%
  filter(winner == 'Away') %>%
  group_by(home_team) %>%
  summarize(home.l.marg = mean(runs.h - runs.a)) %>%
  rename('team' = 'home_team')
margins.away.wins <- team.game.logs %>%
  filter(winner == 'Away') %>%
  group_by(away_team) %>%
  summarize(away.w.marg = mean(runs.a - runs.h)) %>%
  rename('team' = 'away_team')
margins.away.loss <- team.game.logs %>%
  filter(winner == 'Home') %>%
  group_by(away_team) %>%
  summarize(away.l.marg = mean(runs.a - runs.h)) %>%
  rename('team' = 'away_team')
margins <- margins.home.wins %>%
  left_join(margins.home.loss, by = 'team') %>%
  left_join(margins.away.wins, by = 'team') %>%
  left_join(margins.away.loss, by = 'team') 
#%>%
  mutate(`home.w/l.diff` = home.w.marg + home.l.marg, 
         `away.w/l.diff` = away.w.marg + away.l.marg)
```

### [Functions]{.underline}

Useful (and elementary) functions that I created to get practice and
make things slightly easier.

#### Pitcher and Batter matchup functions

Input the teams playing and get comparison stats. Also adds the league
means for context. Don't need to update this every day.

```{r}
batting.matchup <- function(away.team, home.team) {
  view(team.batting %>%
         filter(Team == away.team | Team == home.team | Team == 'Mean'))
}     

pitching.matchup <- function(away.team, home.team) {
  view(team.pitching %>%
         filter(Team == away.team | Team == home.team | Team == 'Mean'))
}
```

#### Individual player game logs

I have provided the Fangraphs call for player game logs AND the function
to make life easier. The Fangraphs call needs a player ID, but the
function just asks for last name, first name, and year.

```{r}
batter.game.logs <- function(last, first, year) {
  x <- playerid_lookup(last, first)
  y <- fg_batter_game_logs(x$fangraphs_id, year)
  view(y)
}
pitcher.game.logs <- function(last, first, year) {
  x <- playerid_lookup(last, first) 
  y <- fg_pitcher_game_logs(x$fangraphs_id, year)
  view(y)
}

Tuve <- fg_batter_game_logs(5417, 2022) %>%
  select(PlayerName, Date, Opp, BatOrder, PA, AB, H, "1B", "2B", "3B", HR, R, RBI, BB, SO, GDP, SB, CS, AVG, GB, FB, LD, 
         Pitches, Balls, Strikes, "Swing%", Pull, Cent, Oppo, EV, "Barrel%", "HardHit%", everything())
x <- fg_pitcher_game_logs(21690, 2023) 
  select(PlayerName, Date, Opp, HomeAway, W, L, ERA, G, GS, IP, TBF, H, R, ER, HR, SO, BB, "H/9", "K/9", "BB/9", "HR/9", 
         WHIP, Balls, Strikes, Pitches, BABIP, "GB%", "FB%", "Swing%", "Barrel%", everything())


```

#### Regression function

Here is a great way to visualize a linear regression between variables.
The function provides you with the R2, intercept, and slope. This is
perfect for a look at whether or not variables have a relationship,
especially before digging deeper.

```{r}
ggplotRegression <- function (fit) {
  require(ggplot2)
  ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) + 
    geom_point() +
    stat_smooth(method = "lm", col = "red") +
    labs(title = paste("Adj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                       "Intercept =",signif(fit$coef[[1]],5 ),
                       " Slope =",signif(fit$coef[[2]], 5),
                       " P =",signif(summary(fit)$coef[2,4], 5)))
}
#example
ggplotRegression(lm(runs.allowed ~ walks.hbp.error, data = f5.season))
```

#### First Five Innings team matchup function (f5.matchup)

This takes a teams home/away splits through the first five innings and
spits out a projected total runs for a matchup. Be sure to put the away
team before the home team. Need to find a way to include pitcher stats
for weight here.

```{r}
f5.matchup <- function(away, home) {
x <- away.five %>%
  filter(away_team == away) %>%
  select(away_team, RS.per.5, RA.per.5, runs.per) %>%
  rename('team' = 'away_team')
y <- home.five %>%
  filter(home_team == home) %>%
  select(home_team, RS.per.5, RA.per.5, runs.per) %>%
  rename('team' = 'home_team')

total <- rbind(x, y) 
print(total %>%
       mutate(f5.score = mean(runs.per)))
}

f5.matchup("TEX", "OAK")
```

#### Predicted Total Score function (total.score)

Takes the same inputs as `f5.matchup` but for whole game instead of five
innings. This is strictly for looking at over/under lines. Does not
include weights for starting pitching which, obviously, it should (and
many other weights too. Baby steps...)

```{r}
total.score <- function(away, home) {
  x <- team.game.logs %>%
    filter(away_team == away) %>%
    group_by(away_team) %>%
    summarize(R.per.game = sum(runs.a)/n(), 
              RA.per.game = sum(runs.h)/n()) %>%
    rename('team' = 'away_team')
  y <- team.game.logs %>%
    filter(home_team == home) %>%
    group_by(home_team) %>%
    summarize(R.per.game = sum(runs.h)/n(), 
              RA.per.game = sum(runs.a)/n()) %>%
    rename('team' = 'home_team')

  
  
  new <- rbind(x, y) %>%
    mutate(ind.score = (R.per.game + RA.per.game)/2) %>%
    mutate(predicted.total = sum(ind.score))

  print(new)
}
```

#### Strength Matchup function (strength.matchup)

This is a simple formula for predicting winners of games found in the
book 'Curveball.' It is meant to be built on. I need to factor in
pitcher performance in their last 3-5 starts as well. Also need to bind
the day's slate of games through an already existing function (but
where?) and then bind this with `team.game.logs` to keep track of the
accuracy of prediction. Note: negative score in away games is good.
Value of 1 is given to a home win, -1 for an away win. Keep that in mind
when assessing the matchup rating.

```{r}
strength.home <- team.game.logs %>%
  group_by(home_team) %>%
  summarize(home.strength = mean(win)) %>%
  rename("team" = "home_team") 
strength.away <- team.game.logs %>%
  group_by(away_team) %>%
  summarize(away.strength = mean(win)) %>%
  rename("team" = "away_team") 
strength <- left_join(strength.home, strength.away, by = 'team') 


strength.matchup <- function(away, home){
  x <- strength %>%
    filter(team == away | team == home) %>%
    arrange(factor(team, levels = away))
  new <- as.data.frame((x[1,3] + x[2,2]))  
  x <- bind_rows(x, new) 

  #print(x)
}

strength.matchup("ATL", "TOR")
```

#### Combined Total Score and Strength function (matchup)

This takes the two previous functions (`total.score` &
`strength.matchup`) and returns a combined tibble. Note: in order to
only return the combined tibble we needed to disable `print` in the
previous two functions. This ensures that it only returns the combined
tibble instead of the other two AND the combined.

```{r}
matchup <- function(away, home){
  x <- strength.matchup(away, home)
  y <- total.score(away, home)
  
  r <- left_join(x, y, by = 'team')
  print(r)
}

matchup("MIN", "LAD")
```

#### Rolling average function

...

### [**Models and Visualization**]{.underline}

#### Predict runs allowed by pitcher

...
